---
title: "Temperature Variation in Sepsis"
author: "Katherine Winner"
date: "`r Sys.Date()`"
output: html_document
---

# I. Set Up

## A. Load Libraries

```{r packages, include = FALSE}
packages <- c("tidyverse", # data manipulation, plotting, etc.
              "lubridate", # simplify dates and times
              "suncalc", # calculate sun and moon positions, phases, etc.
              "comorbidity", # easy comorbidity table based on ICD-10 codes
              "tableone", # descriptive statistics tables
              "patchwork", # combine multiple ggplot2 plots
              "broom", # tidy regression data frames
              "margins", # calculate marginal effects
              "here", # reproducible file paths relative to project root
              "data.table", 
              "arrow", # read/write Parquet files
              "rprojroot",
              "yaml")

install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}
sapply(packages, install_if_missing)
rm(packages, install_if_missing)

# use dplyr select as default
select <- dplyr::select
```

## B. Config and Setting

***load config.yaml*** OR ***Enter site details manually***

```{r config}
# set working directory as project root
project_root <- file.path(find_root(rprojroot::has_dir("Temperature_Variation_Suspected_Sepsis")),
                          "Temperature_Variation_Suspected_Sepsis")
knitr::opts_knit$set(root.dir = project_root)
getwd()

# load config.yaml
config = yaml::read_yaml(paste0(project_root, '/config.yaml'))
### user entered site details
site = tolower(config$site)
file_type = tolower(config$file_type)
tables_location = config$tables_location
project_location = config$project_location
lon = config$lon
lat = config$lat

# ### user entered site details
# site = "Michigan" # add local site name
# file_type = "parquet" # options are "parquet" or "csv"
# tables_location = "Z:/Temperature_CLIF/icu-clif_data_processed" # add local path to CLIF tables
# project_location = "Z:/Temperature_CLIF" # add local path to project location, where results will output
# # Site location details
# lon <- -83.7296 # specify longitude (-122.6878 for OHSU)
# lat <- 42.2838 # specify latitude (45.4996 for OHSU)
# 
# knitr::opts_knit$set(root.dir = project_location)
# getwd()

# create  sub folders within project folder
# check if the output directory exists; if not, create it
if (!dir.exists(paste0(project_location, "/project_tables"))) {
  dir.create(paste0(project_location, "/project_tables"))
}
project_tables <- file.path(paste0(project_location, "/project_tables"))
                            
if (!dir.exists(paste0(project_location, "/project_figures"))) {
  dir.create(paste0(project_location, "/project_figures"))
}
project_figures <- file.path(paste0(project_location, "/project_figures"))

# sink() diverts all standard R outputs to a specified file
# call sink() with filename
sink(file.path(project_tables, "consort_numbers.txt"))
```

## C. Required CLIF Tables

```{r required_tables}
# list of table names from CLIF 2.1
tables <- c("adt", "code_status", "crrt_therapy", "ecmo", "hospital_diagnosis", "hospitalization", "ino", "labs", "medication_admin_continuous", "medication_admin_intermittent", "microbiology_culture", "microbiology_nonculture", "microbiology_susceptibility", "patient", "patient_assessments", "respiratory_support", "vitals")

# tables that should be set to TRUE for this project
true_tables <- c("patient", "hospitalization", "hospital_diagnosis", "adt", "vitals", "labs", "medication_admin_intermittent", "medication_admin_continuous", "microbiology_culture", "patient_assessments", "respiratory_support")

# create a named vector and set the boolean values
table_flags <- setNames(tables %in% true_tables, tables)
```

### 1. Load Tables

```{r load_tables}
# list all CLIF files in the directory
clif_table_filenames <- list.files(path = tables_location,
                                   pattern = paste0("^clif_.*\\.", file_type, "$"),
                                   full.names = TRUE)

# extract the base names of the files (without extension)
clif_table_basenames <- basename(clif_table_filenames) |>
  str_remove(paste0("\\.", file_type, "$"))

# create a lookup table for required files based on table_flags
required_files <- paste0("clif_", names(table_flags) [table_flags])

# check if all required files are present
missing_tables <- setdiff(required_files, clif_table_basenames)
if (length(missing_tables) > 0) {
  stop(paste("Error: Missing required tables:", paste(missing_tables, collapse = ", ")))
}

# filter only by the filenames that are required
required_filenames <- clif_table_filenames[clif_table_basenames %in% required_files]

cast_large_utf8_to_utf8 <- function(x) {
  # Only applies to Arrow objects that have a schema()
  # open_dataset() returns an Arrow Dataset / query
  sch <- arrow::schema(x)

  large_cols <- vapply(
    sch$fields,
    function(f) f$type$ToString() %in% c("large_string", "large_utf8"),
    logical(1)
  )

  cols_to_cast <- sch$names[large_cols]
  if (length(cols_to_cast) == 0) return(x)

  x %>% mutate(across(all_of(cols_to_cast), ~ arrow::cast(.x, arrow::utf8())))
}

# read the required files into a list of data frames
if (file_type == "parquet") {
  data_list <- lapply(required_filenames, open_dataset)
} else if (file_type == "csv") {
  data_list <- lapply(required_filenames, read_csv)
} else if (file_type == "fst") {
  data_list <- lapply(required_filenames, read.fst)
} else {
  stop("Unsupported file format")
}

# assign the data frames to variables based on their file names
for (i in seq_along(required_filenames)) {
  # extract the base name of the file (without extension)
  object_name <- str_remove(basename(required_filenames[i]), paste0("\\.", file_type, "$"))
  # make the object name valid for R (replace invalid characters with underscores)
  object_name <- make.names(object_name)
  # assign the tibble to a variable with the name of the file
  assign(object_name, data_list[[i]])
}
```

## D. Inclusion criteria
Identify hospitalizations for adults >= 18 who were ever in an ED, Ward or ICU 

```{r inclusion}
# keep ward, ICU, and ER encounters (in case of ED to ICU within one system)
inpatient_hospitalization_ids <- clif_adt %>%
  filter(tolower(location_category) %in% c("ward", "icu", "ed", "stepdown")) %>%
  select(hospitalization_id) %>%
  collect()

clif_hospitalization <- clif_hospitalization %>%
  collect() %>% # have to bring in environment to correctly filter IDs
  filter(
    # only keep hospitalizations with locations of interest
    hospitalization_id %in% inpatient_hospitalization_ids$hospitalization_id,
    # only keep adults (no pediatric patients)
    age_at_admission >= 18) %>%
  as_arrow_table() # put back to arrow table
rm(inpatient_hospitalization_ids)

# keep track for consort diagram
patients <- length(unique(clif_hospitalization$patient_id))
encounters <- length(clif_hospitalization$hospitalization_id)
cat('\n In', site, 'CLIF data there are', patients, 'unique patients with', encounters, 'encounters \n')

study_start <- min(clif_hospitalization$admission_dttm)
study_end <- max(clif_hospitalization$admission_dttm)
print(paste("The study period starts on", study_start, "and ends on", study_end))
```

# II. Suspected Infection

## A. Identify Suspected Infection
Use modified CLIF_sepsis code to identify hospitalizations with suspected infection
https://github.com/Common-Longitudinal-ICU-data-Format/CLIF_sepsis/tree/main

Identify hospitalizations with cultures obtained and qualifying antibiotics initiated ***by 24 hours after admission***

### 1. Blood Culture Obtained
(irrespective of the result)

```{r blood_culture}
# Step 1: Load and filter cultures data with specific columns, defining blood_culture_day
blood_cultures <- clif_microbiology_culture %>%
  filter(fluid_category == "blood/buffy coat") %>%
  select(hospitalization_id, collect_dttm, fluid_category) %>%
  rename(blood_culture_collect_time = collect_dttm) %>%
  distinct() %>%
  collect() %>%
  left_join(clif_hospitalization %>%
              select(patient_id, hospitalization_id, admission_dttm) %>%
              collect(),
            by = "hospitalization_id") %>%
  mutate(blood_culture_collect_time = as.POSIXct(blood_culture_collect_time),
         admission_dttm = as.POSIXct(admission_dttm),
         culture_within_24hr = (blood_culture_collect_time - admission_dttm) <= as.difftime(24, units = "hours")) %>%
  filter(culture_within_24hr) %>%
  select(patient_id, hospitalization_id, blood_culture_collect_time, admission_dttm)

# keep track for consort diagram
patients <- length(unique(blood_cultures$patient_id))
encounters <- length(unique(blood_cultures$hospitalization_id))
cat('\n In', site, 'CLIF data there are', patients, 'unique patients with', encounters, 'encounters with blood cultures collected \n')
```

### 2. Qualifying Antibiotics Started

```{r QAD}
# Step 2: Load and filter qualifying antibiotic days, keeping times in exact datetime format
qualifying_abx <- clif_medication_admin_intermittent %>%
  filter(med_group == "CMS_sepsis_qualifying_antibiotics") %>%
  select(hospitalization_id, admin_dttm) %>%
  collect() %>%
  rename(abx_dttm = admin_dttm) %>%
  distinct() %>%
  left_join(clif_hospitalization %>%
              select(patient_id, hospitalization_id, admission_dttm) %>%
              collect(),
            by = "hospitalization_id") %>%
  mutate(abx_dttm = as.POSIXct(abx_dttm),
         admission_dttm = as.POSIXct(admission_dttm),
         abx_within_24hr = (abx_dttm - admission_dttm) <= as.difftime(24, units = "hours")) %>%
  filter(abx_within_24hr) %>%
  select(patient_id, hospitalization_id, abx_dttm, admission_dttm)

# keep track for consort diagram
patients <- length(unique(qualifying_abx$patient_id))
encounters <- length(unique(qualifying_abx$hospitalization_id))
cat('\n In', site, 'CLIF data there are', patients, 'unique patients with', encounters, 'encounters with qualifying antibiotics initiated \n')
```

### 3. Final Suspected Infection 

```{r suspected_infection}
suspected_infection_table <- blood_cultures %>%
  inner_join(qualifying_abx %>%
               select(patient_id, hospitalization_id),
             by = c("patient_id", "hospitalization_id")) %>%
  distinct(hospitalization_id, .keep_all = T)

# keep track for consort diagram
patients <- length(unique(suspected_infection_table$patient_id))
encounters <- length(unique(suspected_infection_table$hospitalization_id))
cat('\n In', site, 'CLIF data there are', patients, 'unique patients with', encounters, 'encounters with suspected infection \n')

# unique encounters with suspected infection
suspected_infection <- suspected_infection_table %>%
  pull(hospitalization_id)
```

# III. Defined Final Cohort

## A. Define Variables

### 1. SOFA Score
SOFA score is calculated in stage 1 using the clifpy co.compute_sofa_scores() command in Python
- In stage 1 we gathered the most extreme SOFA calculated within the first 24hrs per hospitalization_id

```{r sofa_scores}
# read sofa scores from stage 1
sofa_24hr <- read_parquet(paste0(tables_location,"/sofa_24hr.", file_type)) %>%
  select(hospitalization_id, sofa_total) %>%
  collect()
```

### 2. Charlson Comorbidity Index
Charlson Comorbidity Index is calculated in Stage 2 using the clifpy calculate_cci() command in python

```{r charlson}
cci <- clif_hospital_diagnosis %>%
  filter(hospitalization_id %in% suspected_infection) %>%
  collect() %>%
  comorbidity(id = "hospitalization_id", code = "diagnosis_code", map = "charlson_icd10_quan", assign0 = FALSE) %>%
  mutate(cci_score = score(., weights = "quan", assign0 = FALSE)) %>%
  mutate(diabetes = as.integer(diab == 1 | diabwc == 1),
         cirrhosis = as.integer(mld == 1 | msld == 1),
         cancer = as.integer(metacanc == 1 | canc == 1)) %>%
  select(hospitalization_id, diabetes, cirrhosis, cancer, rend, cci_score)
```

### 3. Culture Positive Pathogen

```{r pathogen}
cultures <- clif_microbiology_culture %>%
  filter(hospitalization_id %in% suspected_infection) %>%
  collect() %>%
  mutate(organism_category = case_when(organism_name == "Pseudomonas aeruginosa ( pseaer )" | organism_name == "Pseudomonas aeruginosa (mucoid variety) ( psam )" ~ "pseudomonas_aeruginosa",
                                       organism_name == "Staphylococcus epidermidis (Coag. Negative Staphylococcus) ( staepi )" | organism_name == "coagulase negative staphylococcus ( staneg )" ~ "coag_negative_staphylococcus", 
                                       organism_name == "Staphylococcus aureus ( staaur )" ~ "staphylococcus_aureus", 
                                       organism_name == "Escherichia coli ( esccol )" ~ "eschericha_coli", 
                                       organism_name == "Enterococcus faecalis ( strfae )" | organism_name == "Enterococcus faecium ( strfac )" ~ "enterococcus",
                                       organism_name == "Klebsiella pneumoniae ( klpnsp )" | organism_name == "Klebsiella pneumoniae ( klepne )" ~ "klebsiella_pneumoniae"))
cultures <- cultures %>%
  filter(!is.na(organism_category))

# format cultures wider
pathogens <- cultures %>%
  distinct(hospitalization_id, organism_category) %>%
  mutate(positive = 1) %>%
  pivot_wider(id_cols = hospitalization_id,
              names_from = organism_category,
              values_from = positive,
              values_fill = 0)
```

### 4. ICU Code

```{r icu}
icu <- clif_adt %>%
  select(hospitalization_id, location_name, location_category) %>%
  filter(hospitalization_id %in% suspected_infection,
         location_category == "icu") %>%
  distinct(hospitalization_id, .keep_all = T) %>%
  collect() %>%
  select(-location_category) %>%
  rename(icu = "location_name")
```

### 5. Neutropenia

```{r neutropenia}
neutropenia <- clif_labs %>%
  select(hospitalization_id:lab_result_dttm, lab_category:reference_unit, lab_speciman_category) %>%
  filter(lab_category == "neutrophils_absolute",
         hospitalization_id %in% suspected_infection) %>%
  collect() %>%
  left_join(clif_hospitalization %>%
              select(hospitalization_id, admission_dttm) %>%
              collect(),
            by = "hospitalization_id") %>%
  mutate(lab_collect_dttm = as.POSIXct(lab_collect_dttm),
         admission_dttm = as.POSIXct(admission_dttm),
         # time from admission to lab collection
         time_to_admit = abs(as.numeric(difftime(lab_collect_dttm, admission_dttm, units = "hours")))) %>% 
  # get first lab (collected closest to admission_dttm)
  group_by(hospitalization_id) %>%
  slice_min(order_by = time_to_admit, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  filter(
    # only keep labs collected within 48 hours of admission
    time_to_admit < 48,
    # only keep hospitalization_id with neutropenia (<= 0.5ul absolute neutrophil count)
    lab_value_numeric <= 0.5) %>%
  pull(hospitalization_id)
```

### 6. Anti-pyretic Medications

```{r anti-pyretics}
anti_pyretic_meds <- tolower(c("Acetminophen","Ibuprofen","Ketorolac","Aspirin","Indomethacin","Prednisone","Methylprednisolone","Prednisolone","Infliximab","Adalimumab","Etanercept","Anakinra","Tocilizumab","Interferon alpha","Interferon beta","Rituximab","Interleukin-2"))

# define anti-pyretic medications
# need to load appropriate mediation tables
anti_pyretics <- clif_medication_admin_intermittent %>%
  select(hospitalization_id, admin_dttm, med_category, med_group) %>%
  filter(med_category %in% anti_pyretic_meds,
         hospitalization_id %in% suspected_infection) %>%
  collect() %>%
  full_join(clif_medication_admin_continuous %>%
              select(hospitalization_id, admin_dttm, med_category, med_group) %>%
              filter(med_group == "vasoactives", 
                     hospitalization_id %in% suspected_infection) %>%
              collect()) %>%
  left_join(clif_hospitalization %>%
              select(hospitalization_id, admission_dttm) %>%
              filter(hospitalization_id %in% suspected_infection) %>%
              collect(),
            by = "hospitalization_id") %>%
  mutate(admin_dttm = as.POSIXct(admin_dttm),
         admission_dttm = as.POSIXct(admission_dttm),
         # get time from admit to medication admin
         time_to_admit = abs(as.numeric(difftime(admin_dttm, admission_dttm, units = "hours")))) %>%
  # only keep medications administered within first 48 hours
  filter(time_to_admit < 48) %>%
  pull(hospitalization_id)
```

### 7. Temperatures

```{r temps}
# get temperature measurements 
temperature_celsius <- clif_vitals %>%
  select(hospitalization_id:vital_name, vital_value) %>%
  filter(hospitalization_id %in% suspected_infection,
         vital_name == "temp_c",
         # only keep values within normal physiologic range
         vital_value > 32 & vital_value < 44) %>%
  collect() %>%
  mutate(vital_dttm = as.POSIXct(str_replace(recorded_dttm, "\\..*$", ""), 
                                     format = "%Y-%m-%d %H:%M:%S", tz = "UTC")) %>%
  rename(temp = "vital_value")

# admission temperature (closest to admission_dttm)
first_temps <- temperature_celsius %>%
  left_join(clif_hospitalization %>%
              select(hospitalization_id, admission_dttm) %>%
              filter(hospitalization_id %in% suspected_infection) %>%
              collect(),
            by = "hospitalization_id") %>%
  mutate(admission_dttm = as.POSIXct(admission_dttm),
         time_to_admit = abs(as.numeric(difftime(vital_dttm, admission_dttm, units = "hours")))) %>%
  # get first temperature (recorded closest to admission_dttm)
  group_by(hospitalization_id) %>%
  slice_min(order_by = time_to_admit, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  rename(admit_temp = "temp")
```

## B. Temperature Cohort

```{r temp_cohort}
# define hospitalization days
all_temps_cohort <- temperature_celsius %>%
  left_join(clif_hospitalization %>%
              filter(hospitalization_id %in% suspected_infection) %>%
              collect(), 
            by = c("hospitalization_id")) %>%
  # define days
  mutate(
    # datetime variable
    admission_dttm = as.POSIXct(admission_dttm, tz = "UTC"),
    # 24 hour day (day in 24 hour intervals from admission_dttm)
    day_24hr = as.integer(floor(difftime(vital_dttm, admission_dttm, units = "days")) + 1),
    date = as.Date(vital_dttm)) %>%
  # hospital day (calendar days from admission_dttm)
  group_by(hospitalization_id) %>%
  mutate(hosp_day = as.integer(date - min(date) + 1)) %>%
  ungroup()
```

### 1. Exclusion Criteria

#### a. 48 Hours
Only keep hospitalization that lasted two full days

```{r 48 hours}
days_2 <- clif_hospitalization %>%
  select(hospitalization_id, admission_dttm, discharge_dttm) %>%
  filter(hospitalization_id %in% suspected_infection) %>%
  collect() %>%
  mutate(admission_dttm = as.POSIXct(admission_dttm),
         discharge_dttm = as.POSIXct(discharge_dttm),
         time_to_discharge = difftime(discharge_dttm, admission_dttm, unit = "hours")) %>%
  filter(time_to_discharge >= 48) %>%
  pull(hospitalization_id)

# keep track for consort diagram
encounters <- length(unique(days_2))
cat('\n In', site, 'CLIF data there are', encounters, 'encounters hospitalized for 48+ hours \n')
```

#### b. 2+ Temperature Measurements 
Only keep hospitalizations with two or more temperatures measured 24-48 hours after admission (on day 2 of hospitalization)

```{r two_temps}
two_temps <- all_temps_cohort %>%
  filter(hospitalization_id %in% days_2,
         day_24hr == 2) %>%
  group_by(hospitalization_id, day_24hr) %>%
  summarize(num_measurements = n(), .groups = "drop") %>%
  filter(num_measurements > 1) %>%
  pull(hospitalization_id)

encounters <- length(unique(two_temps))
cat('\n In', site, 'CLIF data there are', encounters, 'encounters with at least 2 temperature mesurements on day 2 of hospitalization \n')
```

### 2. Final Cohort

```{r final_cohort}
final_cohort <- all_temps_cohort %>%
  # only keep hospitalizations admitted at least 48 hours with at least 2 temperature measurements
  filter(hospitalization_id %in% two_temps) %>%
  # get first temps
  left_join(first_temps %>%
              select(hospitalization_id, admit_temp),
            by = "hospitalization_id") %>%
  # get patient demographics
  left_join(clif_patient %>%
              select(patient_id, sex_name, death_dttm) %>%
              collect(),
            by = "patient_id") %>%
  # get SOFA score
  left_join(sofa_24hr, by = "hospitalization_id") %>%
  # get comorbidities
  left_join(cci, by = "hospitalization_id") %>%
  # get culture positive pathogens
  left_join(pathogens, by = "hospitalization_id") %>%
  # get location
  left_join(icu, by = "hospitalization_id") %>%
  # define confounding variables
  mutate(
    # admission temperature category
    admit_temp_cat = factor(case_when(admit_temp < 36.0 ~ "hypothermic",
                                      admit_temp >= 38.3 ~ "hyperthermic",
                                      TRUE ~ "normothermic")),
    neutropenia = as.integer(hospitalization_id %in% neutropenia),
    anti_pyretics = as.integer(hospitalization_id %in% anti_pyretics),
    # age categorized by decade
    age_group = cut(age_at_admission, breaks = c(18, 29, 39, 49, 59, 69, 79, Inf),
                         labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80+"), right = F),
    female = as.integer(sex_name == "Female"),
    pathogen = as.integer(pseudomonas_aeruginosa == 1 | staphylococcus_aureus == 1 | eschericha_coli == 1 | klebsiella_pneumoniae == 1 | enterococcus == 1),
    # if culture is not positive for one of the defined pathogens above = 0
    pathogen = ifelse(is.na(pathogen) | pathogen == 0, 0, 1),
    across(pseudomonas_aeruginosa:eschericha_coli, ~ ifelse(is.na(.x) | .x == 0, 0L, 1L)),
    # get time from admission to death for 30-day mortality
    death_dttm = as.POSIXct(death_dttm),
    time_to_death = as.numeric(difftime(death_dttm, admission_dttm, units = "days")),
    time_to_death = ifelse(is.na(time_to_death), 31, time_to_death),
    die_within_30d = as.integer(time_to_death <= 30))
```

### 3. Temperature Variation

#### a. Diurnal Maximum

```{r sunrise_sunset}
final_cohort <- final_cohort %>%
  left_join(
    # extracts unique dates then returns sunrise and sunset times for each date
    getSunlightTimes(date = unique(.$date), lat = lat, lon = lon, keep = c("sunrise", "sunset")) %>%
              select(date, sunrise, sunset),
            by = "date") %>%
  mutate(
    # determine day or night based on sunrise and sunset times by date
    day_night = if_else(vital_dttm >= sunrise & vital_dttm < sunset, "day", "night")) 

# summarize day/night temperature statistics per encounter+day
temp_wide <- final_cohort %>%
  group_by(hospitalization_id, day_24hr, day_night) %>%
  # summarize min and max temps for each day and night per encounter
  summarize(max_temp = max(temp, na.rm = T),
            min_temp = min(temp, na.rm = T),
            .groups = "drop") %>%
  # pivot wider to get day and night temperatures in the same row
  pivot_wider(names_from = day_night, 
              values_from = c(max_temp, min_temp),
              names_sep = "_") %>%
  # calculate the maximum difference between day and night temperatures
  mutate(
    max_difference = pmax(
      abs(max_temp_day - min_temp_night),
      abs(max_temp_day - max_temp_night),
      abs(min_temp_day - min_temp_night),
      abs(min_temp_day - max_temp_night),
      na.rm = T))
```

#### b. Fever
Where fever is defined as a temperature >= 38.3 Celsius
And normothermia is < 38.3 and > 36.0 Celsius

```{r fever}
get_day_fever <- function(df, day_n) {
  # define new column names
  day_str <- paste0("day", day_n)
  fever_col <- paste0(day_str, "_fever")
  normothermic_col <- paste0(day_str, "_normothermic")
  # summarize max, mix, fever, and normothermic for each hospitalization by day
  df %>%
    filter(day_24hr == day_n) %>%
    group_by(hospitalization_id) %>%
    summarize(
      !!fever_col := as.integer(max(temp, na.rm = T) >= 38.3),
      !!normothermic_col := as.integer(max(temp, na.rm = T) <= 38.3 & min(temp, na.rm = T) > 36.0),
      .groups = "drop")
}

# use function to get fever and normothermia for hospital days 1 & 2
day_1 <- get_day_fever(final_cohort, 1)
day_2 <- get_day_fever(final_cohort, 2)
```

#### c. SD and Deltas

```{r var_cohort}
variation_cohort <- final_cohort %>%
  # get maximum day/night temperature difference
  full_join(temp_wide %>%
              select(hospitalization_id, day_24hr, max_difference),
            by = c("hospitalization_id", "day_24hr")) %>%
  # get day 1 and 2 fever/normothermia classifications
  left_join(day_1, by = "hospitalization_id") %>%
  left_join(day_2, by = "hospitalization_id") %>%
  # get temeprature variation measures by day of hospitalization
  group_by(hospitalization_id, day_24hr) %>%
  mutate(
    # temperature standard deviation
    sd_temp = sd(temp),
    # change in temperature from patient mean
    delta_mean = max(abs(temp - mean(temp))),
    # change in temperature from "normal" (37 Celsius)
    delta_normal = max(abs(temp - 37)),
    # change in temperature from fever
    delta_fever = max(abs(temp - 38.3)),
    # number of temeprature measurements
    num_measurements = n()) %>%
  ungroup()
```

#### d. Quartile

```{r quartile}
# get unique cohort (single row per encounter) by hospital day 2 temperatures 
variation_unique <- variation_cohort %>%
  filter(day_24hr == 2) %>%
  group_by(hospitalization_id) %>%
  slice_max(delta_mean, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(delta_mean = round(delta_mean, 2))
# define quartile bins
quartile <- quantile(variation_unique$delta_mean, probs = seq(0, 1, 0.25), na.rm = T)
# get unique variables
variation_unique <- variation_unique %>%
  mutate(mean_cat = cut(
    delta_mean, 
    breaks = quartile, 
    include.lowest = T,
    labels = as.character(1:4)))

mean_delta <- mean(variation_unique$delta_mean, na.rm = T)
min_delta <- min(variation_unique$delta_mean, na.rm = T)
max_delta <- max(variation_unique$delta_mean, na.rm = T)
pdf(file.path(project_figures,"histogram.pdf"), height = 8, width = 8)
# histogram
variation_unique %>%
  ggplot(aes(x = delta_mean, fill = mean_cat)) +
  geom_histogram(binwidth = 0.01) +
  labs(x = "Temperature Variability\n(maximum change from patient mean)", y = "Number of Patients", fill = "Quartile") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black")) +
  geom_vline(xintercept = mean_delta, linetype = "solid") + # vline based on mean(delta_mean)
  annotate("text", 
           x = mean_delta + 0.025, 
           y = max(table(variation_unique$delta_mean))*0.95, 
           label = paste0("Mean: ", round(mean_delta, 2)), # label based on mean(delta_mean)
           hjust = 0, size = 4) + 
  coord_cartesian(xlim = c(0, 3.0),
                  ylim = c(0, 1000)) +
  scale_x_continuous(breaks = seq(0, 3.0, by = 0.25), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("1" = "#004D40", "2" = "#FFC107", "3" = "#1E88E5", "4" = "#D81B60"))
dev.off()

print(paste("The mean temperature variability across the cohort was", mean_delta,"\u00B0C, the minimum temperature variability was", min_delta,"\u00B0C, and the maximum temperature variability was", max_delta,"\u00B0C"))
```

### 4. Table 1
Descriptive statistics by temperature variation quartile

```{r table 1}
# factorize variables
mean_cat_tab <- variation_unique %>%
  mutate(across(c(die_within_30d, female, diabetes, cirrhosis, cancer, rend, pathogen, neutropenia, mean_cat, admit_temp_cat, day1_fever, day2_fever, pathogen, pseudomonas_aeruginosa, enterococcus, staphylococcus_aureus, klebsiella_pneumoniae, eschericha_coli),
                as.factor))
# variables to include
vars <- c("delta_mean", "female", "age_at_admission", "sofa_total", "cci_score", "diabetes", "rend", "cirrhosis", "cancer", "neutropenia", "admit_temp_cat", "day1_fever", "day2_fever", "pathogen", "die_within_30d", "pseudomonas_aeruginosa", "enterococcus", "staphylococcus_aureus", "klebsiella_pneumoniae", "eschericha_coli")
catvars <- c("female", "diabetes", "rend", "cirrhosis", "cancer", "die_within_30d", "neutropenia", "admit_temp_cat", "pathogen", "day1_fever", "day2_fever", "pseudomonas_aeruginosa", "enterococcus", "staphylococcus_aureus", "klebsiella_pneumoniae", "eschericha_coli")
contvars <- setdiff(vars, catvars)

# table stratified by mean_cat
tbl <- CreateTableOne(vars = vars, data = mean_cat_tab, strata = "mean_cat", factorVars = catvars)
tbl <- print(tbl, nonormal = NULL, pval = T, quote = F, noSpaces = T, pDigits = 3, printToggle = F)
tbl_df <- as.data.frame(tbl)
# full cohort table
tbl_ttl <- CreateTableOne(vars = vars, data = mean_cat_tab, factorVars = catvars)
tbl_ttl <- print(tbl_ttl, nonormal = NULL, quote = F, noSpaces = T, printToggle = F)
tbl_ttl_df <- as.data.frame(tbl_ttl)

# get rows that correspond to continuous variables for formatting
cont_rows <- Reduce('|', lapply(contvars, function(v) startsWith(rownames(tbl_df), v)))
# replace x (y)" cell text -> "x ± y" for continuous variables
tbl_df[cont_rows, ] <- lapply(tbl_df[cont_rows, ], function(x) {
  x <- as.character(x)
  gsub("(-?\\d+(?:\\.\\d+)?) \\((\\d+(?:\\.\\d+)?)\\)", "\\1 ± \\2", x)
})

tbl_ttl_df[cont_rows, ] <- lapply(tbl_ttl_df[cont_rows, ], function(x) {
  x <- as.character(x)
  gsub("(-?\\d+(?:\\.\\d+)?) \\((\\d+(?:\\.\\d+)?)\\)", "\\1 ± \\2", x)
})

# write csv
write.csv(tbl_df, file.path(project_tables, "table1_cat.csv"))
write.csv(tbl_ttl_df, file.path(project_tables, "table1_all.csv"))
```

## D. Why Day 2?

### 1. Average temp. by day

```{r why_day2}
# group by day-hour
variation_cohort$hour <- as.integer(format(variation_cohort$vital_dttm, "%H"))
summary_by_hour <- variation_cohort %>%
  filter(day_24hr %in% c(1:6),
         hour %in% seq(0, 20, by = 4)) %>%
  group_by(day_24hr, hour) %>%
  summarize(
    mean_temp = mean(temp, na.rm = T),
    .groups = "drop") %>%
  mutate(day_label = factor(paste0("Day ", day_24hr)))

# plot
pdf(file.path(project_figures,"avg_temp_day.pdf"), height = 8, width = 11)
ggplot(summary_by_hour, aes(hour, mean_temp, color = factor(day_24hr))) +
  geom_line(linewidth = 1.1) +
  scale_x_continuous(breaks = seq(0, 24, by = 4),
                     labels = c("0:00", "4:00", "8:00", "12:00", "16:00", "20:00", "0:00"),
                     expand = c(0,0)) +
  # scale_y_continuous(breaks = seq(36.8, 37.1, by = 0.1)) +
  labs(y = "Mean Temperature (°C)",
       x = "Hour of Day",
       color = "Day") +
  scale_color_manual(values = c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "#FFB000", "#80EC9F")) +
  coord_cartesian(ylim = c(36.8,37.1)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12))
dev.off()
```

### 2. Temperature trajectory

```{r temp_trajectory}
# compute continuous time since admission in hours
hourly_average <- variation_cohort %>%
  filter(day_24hr %in% c(1:6)) %>%
  mutate(continuous_time = (day_24hr - 1) * 24 + hour,
         day = day_24hr,
         day_label = factor(paste0("Day ", day_24hr))) %>%
  group_by(continuous_time, day, day_label) %>%
  summarize(avg_temp = mean(temp, na.rm = T),
            .groups = "drop") %>%
  arrange(continuous_time)
# get coordinates for beginning and end of day for continuous line
hourly_average_segs <- hourly_average %>%
  mutate(next_x = lead(continuous_time),
         next_y = lead(avg_temp),
         next_day = lead(day)) %>%
  filter(!is.na(next_x))
# panel 1: line plot of mean hourly temperature
labels <- rep(c("0:00", "12:00"), 6)
labels <- c(labels, "0:00")
pdf(file.path(project_figures,"temp_trajectory.pdf"), height = 8, width = 11)
ggplot(hourly_average_segs) +
  geom_segment(aes(
    x = continuous_time, y = avg_temp,
    xend = next_x, yend = next_y, # create continuous line
    color = factor(day)),
    size = 1) +
  scale_x_continuous(
    breaks = seq(0, 144, by = 12),
    labels = labels,
    expand = c(0,0)) +
  coord_cartesian(ylim = c(36.8, 37.1)) +
  scale_color_manual(values = c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "#FFB000", "#80EC9F"),
                     name = "Day") +
  labs(y = "Mean Hourly Temperature",
       x = "Hour of Day") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.line = element_line(color = "black"),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12))
dev.off()

temperature_variation <- variation_cohort %>%
  filter(day_24hr %in% c(1:6)) %>%
  group_by(day_24hr, hospitalization_id) %>%
  slice_max(delta_mean, n = 1, with_ties = FALSE) %>%
  ungroup()

# panel 2: temperature variation boxplots
pdf(file.path(project_figures,"temp_variability.pdf"), height = 8, width = 11)
ggplot(temperature_variation, aes(x = factor(day_24hr), y = delta_mean, fill = factor(day_24hr), alpha = 0.08)) +
  geom_boxplot(outlier.shape = NA, coef = 0) +
  scale_fill_manual(values = c("#648FFF", "#785EF0", "#DC267F", "#FE6100", "#FFB000", "#80EC9F")) +
  labs(y = "Temperature Variability",
       x = "Day") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 1), xlim = c(1, 6)) +
  theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12))
dev.off()
```

# IV. Outcomes

## A. 30-day Mortality

```{r mortality}
temp_event <- variation_unique %>%
  mutate(mean_cat4 = as.integer(mean_cat == 4),
         admit_temp_cat  = factor(admit_temp_cat, levels = c("normothermic", "hyperthermic", "hypothermic")),
         age_group = as.numeric(age_group))

## logistic regression of mortality
# create a list of predictors
predictors <- list(
  "delta_mean" = "delta_mean",
  "mean_cat4" = "mean_cat4",
  "sd_temp" = "sd_temp",
  "max_difference" = "max_difference",
  "delta_normal" = "delta_normal",
  "delta_fever" = "delta_fever",
  "day1_fever" = "day1_fever",
  "day2_fever" = "day2_fever")
# create a list of covariates
adj_covars <- c("sofa_total", "age_group", "female", "diabetes", "rend", "cancer", "cirrhosis", "neutropenia", "pathogen", "icu", "anti_pyretics", "num_measurements", "admit_temp_cat")

## unadjusted analysis
# function to fit and tidy logistic regression
fit_logit_model <- function(predictor) {
  formula <- as.formula(paste("die_within_30d ~", predictor))
  glm(formula, data = temp_event, family = binomial) %>%
    broom::tidy(conf.int = TRUE, exponentiate = TRUE)
}
# run unadjusted model on all predictors
logit_results <- map_dfr(
  predictors,
  ~fit_logit_model(.x) %>%
    filter(term == .x | (grepl(.x, term) & .x %in% c("day1_fever", "day2_fever"))), # keep predictor row only
  .id = "predictor"
)

## adjusted analysis
# function to fit and tidy adjusted model
fit_logit_adj <- function(predictor) {
  formula <- as.formula(
    paste("die_within_30d ~", paste(c(predictor, adj_covars), collapse = "+"))
  )
  glm(formula, data = temp_event, family = binomial) %>%
    broom::tidy(conf.int = TRUE, exponentiate = TRUE)
}
# run adjusted model on all predictors
adj_logit_results <- map_dfr(
  predictors, 
  ~fit_logit_adj(.x) %>%
      filter(term == .x | (grepl(.x, term) & .x %in% c("day1_fever", "day2_fever"))), # keep predictor row only
  .id = "predictor"
)

# combine model results
logit_results <- logit_results %>% 
  mutate(model = "Unadjusted")
adj_logit_results <- adj_logit_results %>% 
  mutate(model = "Adjusted")
all_results <- bind_rows(logit_results, adj_logit_results) %>%
  select(model, predictor, term, estimate, conf.low, conf.high, p.value)

write.csv(all_results, file.path(project_tables, "mortality_logit_results.csv"))
```

```{r mortality_forest_plot}
logit_labels <- c('delta_mean' = 'Delta Temperature',
                  'mean_cat4' = '4th Quartile Delta Temperature',
                  'sd_temp' = 'Temperature SD',
                  'max_difference' = 'Diurnal Maximum',
                  'day1_fever' = 'Day 1 Fever',
                  'day2_fever' = 'Day 2 Fever')

pdf(file.path(project_figures,"mortality_logit_results.pdf"), height = 8, width = 8)
adj_logit_results %>%
  mutate(term_label = logit_labels[as.character(term)]) %>%
  filter(!is.na(term_label)) %>%
  mutate(term_label = forcats::fct_reorder(term_label, estimate)) %>%
  ggplot(aes(x = term_label, y = estimate)) +
  geom_point(color = "#D81B60", size = 2) +
  geom_segment(aes(xend = term_label, y = conf.low, yend = conf.high), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = NULL,
    y = "Odds Ratio of 30-Day Mortality (95% CI)") +
  scale_y_continuous(limits = c(0.5, 2.15), breaks = seq(0.5, 2.0, by = 0.5)) +
  coord_flip() +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.title.x = element_text(color = "black", size = 16),
    axis.title.y = element_text(color = "black", size = 16),
    axis.text.x = element_text(color = "black", size = 12),
    axis.text.y = element_text(color = "black", size = 12))
dev.off()
```

### 1. Marginal Effects

```{r margins}
model_covariate <- glm(die_within_30d ~ delta_mean + sofa_total + age_group + female + diabetes + rend + cancer + cirrhosis + neutropenia + icu + anti_pyretics + pathogen + admit_temp_cat + num_measurements, data = temp_event, family = "binomial")
broom::tidy(model_covariate, conf.int = TRUE, exponentiate = TRUE)

# get marginal effects
model_margins <- margins(model_covariate)
tidy_margins <- broom::tidy(model_margins, conf.int = TRUE) %>%
  filter(!str_detect(term, "Intercept"))
write.csv(tidy_margins, file.path(project_tables, "mortality_margins.csv"))

# create variable labels for plotting
variable_labels <- c(
  'delta_mean' = 'Temperature Variability',
  'delta_mean_sensitivity' = 'Temperature Variability',
  'mean_cat4' = '4th quartile temperature var.',
  'mean_cat3' = '3rd quartile temperature var.',
  'mean_cat2' = '2nd quartile temperature var.',
  'admit_temp_cathypothermic' = 'Hypothermic',
  'admit_temp_cathyperthermic' = 'Hyperthermic',
  'cancer' = 'Cancer',
  'normothemic' = 'Normothermic',
  'cirrhosis' = 'Cirrhosis',
  'decade' = 'Age (decade)',
  'sofa_total' = 'SOFA score',
  'rend' = 'End-stage renal disease',
  'female' = 'Female sex',
  'neutropenia' = 'Neutropenia',
  'num_measurements' = '# Temperature measurements',
  'icu' = 'Admission ICU',
  'diabetes' = 'Diabetes',
  'pathogen' = 'Culture positive bacteremia',
  'anti_pyretics' = 'Anti-pyretics')
# # plot marginal effects
# plot_margins <- tidy_margins %>%
#   mutate(term_label = variable_labels[as.character(term)]) %>%
#   filter(!is.na(term_label)) %>%
#   mutate(term_label = forcats::fct_reorder(term_label, estimate)) %>%
#   ggplot(aes(x = term_label, y = estimate)) +
#   geom_segment(aes(xend = term_label, y = conf.low, yend = conf.high)) +
#   geom_point(color = "#D81B60", size = 2) +
#   geom_hline(yintercept = 0, linetype = "dashed") +
#   labs(x = NULL, y = "Average Marginal Effect") +
#   coord_flip() +
#   theme_minimal(base_size = 14) +
#   theme(
#     legend.position = "none",
#     panel.grid.major = element_blank(),
#     panel.grid.minor = element_blank(),
#     axis.line = element_line(color = "black"),
#     axis.title.x = element_text(color = "black", size = 16),
#     axis.title.y = element_text(color = "black", size = 16),
#     axis.text.x = element_text(color = "black", size = 12),
#     axis.text.y = element_text(color = "black", size = 12))
# print(plot_margins)
```

### 2. Measurement Bias - sensitivity analysis

```{r measurement_bias}
# for each hour, calculate how often a temperature was recorded and how many were abnormal
abnormal_counts <- variation_cohort %>%
  # create a binary indicator for abnormal temperature
  mutate(abnormal_temp = as.integer(temp < 36.0 | temp >= 38.3)) %>% 
  group_by(hour) %>%
  summarize(hourly_count = n(),
            abnormal_count = sum(abnormal_temp),
            # calculate probability of temperature being abnormal
            probability_abnormal = abnormal_count / hourly_count,
            .groups = "drop") %>%
  mutate(total_count = sum(hourly_count),
         # calculate probability that temperature will be measured at a given hour
         probability_measurement = hourly_count / total_count)

# get routine hours defined as times when probability of  measurement is highest and probability of abnormal temperature is lowest
# want 6 times spaced 4 hours apart
# generate 4 hour sets
candidate_hours <- lapply(0:3, function(start) { 
  routine_hours <- seq(start, 23, by = 4)
  set_stats <- abnormal_counts %>%
    filter(hour %in% routine_hours) %>%
    summarize(
      sum_measurement = sum(probability_measurement),
      sum_abnormal = sum(probability_abnormal)
    )
  # evaluate each set
  tibble(
    start_hour = start,
    hours = list(routine_hours),
    sum_measurement = set_stats$sum_measurement,
    sum_abnormal = set_stats$sum_abnormal
  )
}) %>%
  bind_rows()

# pick optimal set based on highest measurement probability & lowest abnormal probability
top_set <- candidate_hours %>%
  arrange(desc(sum_measurement), sum_abnormal) %>%
  slice(1)

routine_hours <- unlist(top_set$hours)

# plot
pdf(file.path(project_figures,"routine_temps.pdf"), height = 8, width = 11)
# Base plot: Bar plot for number of temperature measurements
ggplot(abnormal_counts, aes(x = hour)) +
  # Bar plot layer
  geom_bar(aes(y = probability_measurement), stat = "identity", fill = "#1E88E5", alpha = 0.3) +
  # Line plot layer for probability of abnormal temperature
  geom_line(aes(y = probability_abnormal), color = "#D81B60", linewidth = 1) +
  geom_point(aes(y = probability_abnormal), color = "#D81B60") +
  scale_x_continuous(breaks = seq(0, 23, 1), expand = c(0, 0)) +
  labs(x = "Hour of Day",
       y = "Probability of Measurement") +
  # Secondary y-axis for probability scale
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 0.125),
    sec.axis = sec_axis(~ ., name = "Probability of Abnormal Temperature", labels = scales::percent),
    expand = c(0, 0)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        legend.position = "none")
dev.off()

# routine hours = 0, 4, 8, 12, 16, 20
# restrict cohort for routine hours of measurement
routine_hours <- variation_cohort %>%
  filter(hour %in% routine_hours) %>%
  group_by(hospitalization_id) %>%
  mutate(delta_mean_sensitivity = max(abs(temp - mean(temp)))) %>%
  slice_max(delta_mean_sensitivity, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(admit_temp_cat  = factor(admit_temp_cat, levels = c("normothermic", "hyperthermic", "hypothermic")),
         age_group = as.numeric(age_group))

routine_model <- glm(die_within_30d ~ delta_mean_sensitivity + sofa_total + age_group + female + diabetes + rend + cancer + cirrhosis + neutropenia + icu + anti_pyretics + pathogen + admit_temp_cat, data = routine_hours, family = "binomial")
routine_model_results <- broom::tidy(routine_model, conf.int = TRUE, exponentiate = TRUE)
write.csv(routine_model_results, file.path(project_tables, "sensitivity_routine_temps.csv"))

pdf(file.path(project_figures, "sensitivity_routine_temps.pdf"), height = 8, width = 8)
broom::tidy(routine_model, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(!str_detect(term, "Intercept")) %>%
  mutate(term_label = variable_labels[as.character(term)]) %>%
  filter(!is.na(term_label)) %>%
  mutate(term_label = forcats::fct_reorder(term_label, estimate)) %>%
  ggplot(aes(x = term_label, y = estimate)) +
  geom_segment(aes(xend = term_label, y = conf.low, yend = conf.high)) +
  geom_point(color = "#D81B60", size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(x = NULL, y = "Odds Ratio of 30-day Mortality") +
  scale_y_continuous(limits = c(0.5, 3.5), breaks = seq(0.5, 3.5, by = 0.5)) +
  coord_flip() +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.title.x = element_text(color = "black", size = 16),
    axis.title.y = element_text(color = "black", size = 16),
    axis.text.x = element_text(color = "black", size = 12),
    axis.text.y = element_text(color = "black", size = 12))
dev.off()
```

### 3. Day of Hospitalization - sensitivity analysis

```{r day_24hr}
# Loop over days, fit logistic regression, extract OR and CI
results <- map_dfr(1:6, function(day_num) {
  temp_event_day <- variation_cohort %>%
    filter(day_24hr == day_num) %>%
    group_by(hospitalization_id) %>%
    slice_max(delta_mean, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    filter(num_measurements > 1) %>%
    mutate(admit_temp_cat  = factor(admit_temp_cat, levels = c("normothermic", "hyperthermic", "hypothermic")),
           age_group = as.numeric(age_group))
  
model <- glm(die_within_30d ~ delta_mean + sofa_total + age_group + female + diabetes + rend + cancer + cirrhosis + neutropenia + icu + anti_pyretics + pathogen + admit_temp_cat + num_measurements, 
               data = temp_event_day, 
               family = "binomial")
  
  broom::tidy(model, conf.int = TRUE, exponentiate = TRUE) %>%
    filter(term == "delta_mean") %>%
    mutate(day_24hr = day_num)
})

hospitalization_day <- results %>%
  mutate(day_label = factor(
    paste0("Day ", day_24hr),
    levels = paste0("Day ", 6:1)))
write.csv(hospitalization_day, file.path(project_tables, "sensitivity_hospitalization_day.csv"))

# plot
pdf(file.path(project_figures,"sensitivity_hospitalization_day.pdf"), height = 8, width = 8)
results %>%
  mutate(day_label = factor(
    paste0("Day ", day_24hr),
    levels = paste0("Day ", 6:1)
  )) %>%
  ggplot(aes(x = day_label, y = estimate)) +
  geom_point(color = "#D81B60", size = 2) +
  geom_segment(aes(x = day_label, xend = day_label, y = conf.low, yend = conf.high), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = "24-Hour Day",
    y = "Odds Ratio of 30-Day Mortality (95% CI)") +
  scale_y_continuous(limits = c(0.75, 2.3), breaks = seq(0.75, 2.25, by = 0.25)) +
  coord_flip() +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.title.x = element_text(color = "black", size = 12),
    axis.title.y = element_text(color = "black", size = 12),
    axis.text.x = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", size = 10))
dev.off()
```

### 4. Fever-free - sensitivity analysis

```{r fever_free}
fever <- temp_event %>%
  filter(day1_fever == 1 | day2_fever == 1) %>%
  pull(hospitalization_id)
no_fever_event <- temp_event %>%
  filter(!hospitalization_id %in% fever)

patients <- length(unique(no_fever_event$patient_id))
encounters <- length(no_fever_event$hospitalization_id)
cat('\n In', site, 'CLIF data there are', patients, 'unique patients and', encounters, 'encounters \n without fever within the first 48 hours of hospitalization')

model_covariate <- glm(die_within_30d ~ delta_mean + sofa_total + age_group + female + diabetes + rend + cancer + cirrhosis + neutropenia + icu + anti_pyretics + pathogen + admit_temp_cat + num_measurements, data = no_fever_event, family = "binomial")
sensitivity_fever <- broom::tidy(model_covariate, conf.int = TRUE, exponentiate = TRUE)
write.csv(sensitivity_fever, file.path(project_tables, "sensitivity_fever.csv"))

# plot
pdf(file.path(project_figures,"sensitivity_fever.pdf"), height = 8, width = 8)
broom::tidy(model_covariate, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(!str_detect(term, "Intercept")) %>%
  mutate(term_label = variable_labels[as.character(term)]) %>%
  filter(!is.na(term_label)) %>%
  mutate(term_label = forcats::fct_reorder(term_label, estimate)) %>%
  ggplot(aes(x = term_label, y = estimate)) +
  geom_segment(aes(xend = term_label, y = conf.low, yend = conf.high)) +
  geom_point(color = "#D81B60", size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(x = NULL, y = "Odds Ratio of 30-day Mortality") +
  scale_y_continuous(limits = c(0.0, 4), breaks = seq(0, 4, by = 0.5)) +
  coord_flip() +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.title.x = element_text(color = "black", size = 16),
    axis.title.y = element_text(color = "black", size = 16),
    axis.text.x = element_text(color = "black", size = 12),
    axis.text.y = element_text(color = "black", size = 12))
dev.off()
```

## B. Bacteremia

```{r bacteremia}
# create a list of covariations
bacter_covars <- c("sofa_total", "age_group", "female", "diabetes", "rend", "cancer", "cirrhosis", "neutropenia", "icu", "anti_pyretics", "num_measurements", "admit_temp_cat")

## unadjusted analysis
# function to fit and tidy logistic regression
fit_bacter_model <- function(predictor) {
  formula <- as.formula(paste("pathogen ~", predictor))
  glm(formula, data = temp_event, family = binomial) %>%
    broom::tidy(conf.int = TRUE, exponentiate = TRUE)
}
# run unadjusted model on all predictors
bacter_results <- map_dfr(
  predictors,
  ~fit_bacter_model(.x) %>%
    filter(term == .x | (grepl(.x, term) & .x %in% c("day1_fever", "day2_fever"))), # keep predictor row only
  .id = "predictor"
)

## adjusted analysis
# function to fit and tidy adjusted model
fit_bacter_adj <- function(predictor) {
  formula <- as.formula(
    paste("pathogen ~", paste(c(predictor, bacter_covars), collapse = "+"))
  )
  glm(formula, data = temp_event, family = binomial) %>%
    broom::tidy(conf.int = TRUE, exponentiate = TRUE)
}
# run adjusted model on all predictors
adj_bacter_results <- map_dfr(
  predictors, 
  ~fit_bacter_adj(.x) %>%
      filter(term == .x | (grepl(.x, term) & .x %in% c("day1_fever", "day2_fever"))), # keep predictor row only
  .id = "predictor"
)

# combine model results 
bacter_results <- bacter_results %>% 
  mutate(model = "Unadjusted")
adj_bacter_results <- adj_bacter_results %>% 
  mutate(model = "Adjusted")
all_bacter_results <- bind_rows(bacter_results, adj_bacter_results) %>%
  select(model, predictor, term, estimate, conf.low, conf.high, p.value)
write.csv(all_bacter_results, file.path(project_tables, "bacteremia_logit_results.csv"))

bacter_model_covariate <- glm(pathogen ~ delta_mean + age_group + female + diabetes + rend + cancer + cirrhosis + neutropenia + icu + anti_pyretics + admit_temp_cat + num_measurements, data = temp_event, family = "binomial")
# get marginal effects
bacter_model_margins <- margins(bacter_model_covariate)
tidy_bacter_margins <- broom::tidy(bacter_model_margins, conf.int = TRUE) %>%
  filter(!str_detect(term, "Intercept"))
tidy_bacter_margins
write.csv(tidy_bacter_margins, file.path(project_tables, "bacteremia_margins.csv"))

pdf(file.path(project_figures,"bacteremia_logit_results.pdf"), height = 8, width = 8)
adj_bacter_results %>%
  mutate(term_label = logit_labels[as.character(term)]) %>%
  filter(!is.na(term_label)) %>%
  mutate(term_label = forcats::fct_reorder(term_label, estimate)) %>%
  ggplot(aes(x = term_label, y = estimate)) +
  geom_point(color = "#D81B60", size = 2) +
  geom_segment(aes(xend = term_label, y = conf.low, yend = conf.high), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = NULL,
    y = "Odds Ratio of 30-Day Mortality (95% CI)") +
  scale_y_continuous(limits = c(0.9, 3.0), breaks = seq(1, 3.0 , by = 0.5)) +
  coord_flip() +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.title.x = element_text(color = "black", size = 16),
    axis.title.y = element_text(color = "black", size = 16),
    axis.text.x = element_text(color = "black", size = 12),
    axis.text.y = element_text(color = "black", size = 12))
dev.off()
```

```{r sink}
# call sink() to return output to file specified earlier
sink()
```
